### 2021-03-07 text/draft.tex
dept results

\begin{figure}
\centering
\includegraphics{images/keywords_counted.pdf}
\caption{The distribution of keywords per item in the Edoc sample data
set. The leftmost boxplot shows the distribution of cleaned keywords
(min = 0, Q1 = 4, M = 6, Q3 = 8, max = 87); the other boxplots show the
distribution of cleaned keywords with reconciled QID (min = 0, Q1 = 2, M
= 4, Q3 = 6, max = 80), MeSH ID (min = Q1 = 0, M = 2, Q3 = 4, max = 63),
and YSO ID respectively (min = Q1 = 0, M = 1, Q3 = 2, max = 46). All
four distributions are similarly consistent, but there is a linear
shrinkage of the center leaving MeSH and YSO with potentially too few
descriptors to represent an adequate indexing.}
\end{figure}

@@

The metrics used for the assessment are precision, recall and F1-score.
Precision and recall are standard metrics for indexing quality (e.g.,
Gantert 2016, 197) whereby the F1 score plays are more prominent role in
the assessment of machine indexing (e.g., Suominen 2018, 11--14; Toepfer
and Kempf 2016, 93f.). Of course, there is a host of alternative metrics
(such as indexing consistency, transparency, reproducability) that are
neglected here.

@@

In this section I will discuss the general results of assessing the
performance of the 100 Annif configurations versus the native gold
standard with respect to the Edoc sample data set.

\begin{figure}
\centering
\includegraphics{images/metrics_all_project+abstract.pdf}
\caption{Distribution of weighted F1-scores per class of project of
Annif configuration (left) and per class of text basis of Annif
configuration (right). \texttt{wikidata} (min = 0.239, Q1 = 0.355, M =
0.415, Q3 = 0.492, max = 0.517) outperforms any YSO configuration.
\texttt{yso-bonsai-en}, \texttt{yso-en} and \texttt{yso-fasttext-en} are
on par (min = 0.169, Q1 = 0.218, M = 0.312, Q3 = 0.414, max = 0.496) and
slightly outperformed by the more consistent \texttt{yso-maui-en} (min =
0.171, Q1 = 0.250, M = 0.327, Q3 = 0.407, max = 0.496). Surprisingly,
the performance of configurations based on titles (min = 0.169, Q1 =
0.250, M = 0.345, Q3 = 0.424, max = 0.517) was better than the
performance of configurations based on titles and abstracts (min =
0.169, Q1 = 0.242, M = 0.337, Q3 = 0.424, max = 0.517).}
\end{figure}

Let us first consider overall performance. Here the most remarkable
result is that Wikidata configurations outperformed YSO configurations
(see Figure 6, left). However, the explanation of this effect is not
evident. The two main explanatory hypotheses are 1. that the suggestions
by the Wikidata configurations are more salient, or 2. that the native
gold standard is biased towards Wikidata due to its higher coverage of
QIDs as compared to YSO IDs. By contrast, the Wikidata configurations
are more productive than the YSO configurations, and the reason for this
effect is the higher coverage of QIDs as compared to YSO IDs (see Figure
7). By ``productivity'' I mean that the absolute number of assigned
subject terms. So Wikidata configurations are not only qualitatively but
also quantitatively superior to YSO configurations.

@@

Furthermore, the four YSO configurations are more or less on par: the
\texttt{yso-bonsai-en}, \texttt{yso-en}, and \texttt{yso-fasttext-en}
configurations perform very similarily and are slightly outperformed by
\texttt{yso-maui-en} (see Figure 6, left).

A surprising find is that there was no performance gain achieved by
increasing the text basis from titles to titles and abstracts (see
section X): if anything, the performance for the extended text basis was
slightly worse (see Figure 6, right).

@@

\begin{figure}
\centering
\includegraphics{images/metrics_all_n.pdf}
\caption{Distribution of weighted F1-scores per maximum number of
suggestions 0 \textless{} n \textless11.}
\end{figure}

we can then discuss whether the best performance is actually useful;
perhaps do this by comparing absolute scores with other assessments. if
results are much lower it is to be suspected that the native gold
standard is not very good (i.e., that the measure we use to judge annif
performance is inadequte =\textgreater{} motivation for foreign
standard)

construct all metrics: \_Analysis.super\_make\_metrics(file\_path=DIR +
``/indexed/indexed\_master\_mesh\_enriched.json'')

construcht all stats for metrics: \_Analysis.super\_make\_stats()

\hypertarget{departments}{%
\subsubsection{Departments}\label{departments}}

I had noted in section X as caveat that the performance of Annif might
vary according to department due to systematic biases in constructing
the Edoc sample data set. It might therefore be illuminating to assess
the performance of the various Annif configurations per department.
Since the Edoc \texttt{department} data field is mandatory, the sample
data set is partioned into blocks of departments by default. We can thus
create a data foundation for each such block by calling
\texttt{\_Analysis.super\_make\_metrics} with a departmental parameter:

@@

This yields 1000 files in
\texttt{/metrics/metrics\_\{department\}\_\{marker\}.json} (100
configurations \(*\) 10 departments) where \texttt{department} specifies
the department according to the Edoc convention.

Let us now look at the results. For each department, we are interested
in the highest weighted F1-score and the corresponding configuration as
summarized in Figure X.

@@

\hypertarget{yso-configurations}{%
\subsubsection{YSO configurations}\label{yso-configurations}}

\hypertarget{wikidata-configurations}{%
\subsubsection{Wikidata configurations}\label{wikidata-configurations}}

@@

\hypertarget{digression-truth-conditions-for-indexing}{%
\subsection{Digression: truth-conditions for
indexing}\label{digression-truth-conditions-for-indexing}}

(Correspondence theory of indexing: subject term T fits text X iff text
X is about T) \#\# Targets !! Find out what I meant here.

\hypertarget{analysis-2}{%
\subsection{Analysis}\label{analysis-2}}

\hypertarget{discussion-1}{%
\subsection{Discussion}\label{discussion-1}}

@@

@@

\leavevmode\hypertarget{ref-Suominen.2018}{}%
Suominen, Osma. 2018. {``Annif: DIY Automated Subject Indexing Using
Multiple Algorithms.''} \emph{LIBER Quarterly} 29 (1): 1--25.
\url{https://doi.org/10.18352/lq.10285}.

>>>>

\begin{figure}
\centering
\includegraphics{images/keywords_counted.pdf}
\caption{The distribution of keywords per item in the Edoc sample data
set. The leftmost boxplot shows the distribution of cleaned keywords
(min = 0, Q1 = 4, M = 6, Q3 = 8, max = 87); the other boxplots show the
distribution of cleaned keywords with reconciled QID (min = 0, Q1 = 2, M
= 4, Q3 = 6, max = 80), MeSH ID (min = Q1 = 0, M = 2, Q3 = 4, max = 63),
and YSO ID respectively (min = Q1 = 0, M = 1, Q3 = 2, max = 46). All
four distributions are similarly consistent, but there is a linear
decrease of the distribution's center leaving MeSH and YSO with
potentially too few descriptors to represent an adequate indexing.}
\end{figure}

@@

The metrics used for the assessment are precision, recall and F1-score.
Precision and recall are standard metrics for indexing quality (e.g.,
Gantert 2016, 197) whereby the F1 score plays are more prominent role in
the assessment of machine indexing (e.g., Suominen 2019, 11--14; Toepfer
and Kempf 2016, 93f.). Of course, there is a host of alternative metrics
(such as indexing consistency, transparency, reproducability) that are
neglected here.

@@

In this section I will discuss the general results of assessing the
performance of the 100 Annif configurations versus the native gold
standard with respect to the Edoc sample data set. The data foundation
is available at \texttt{/analysis/metrics.json}.

\begin{figure}
\centering
\includegraphics{images/metrics_all_project+abstract.pdf}
\caption{Distribution of weighted F1-scores per class of project of
Annif configuration (left) and per class of text basis of Annif
configuration (right). \texttt{wikidata} (min = 0.239, Q1 = 0.355, M =
0.415, Q3 = 0.492, max = 0.517) outperforms any YSO configuration.
\texttt{yso-bonsai-en}, \texttt{yso-en} and \texttt{yso-fasttext-en} are
on par (min = 0.169, Q1 = 0.218, M = 0.312, Q3 = 0.414, max = 0.496) and
slightly outperformed by the more consistent \texttt{yso-maui-en} (min =
0.171, Q1 = 0.250, M = 0.327, Q3 = 0.407, max = 0.496). Surprisingly,
the performance of configurations based on titles (min = 0.169, Q1 =
0.250, M = 0.345, Q3 = 0.424, max = 0.517) was marginally better than
the performance of configurations based on titles and abstracts (min =
0.169, Q1 = 0.242, M = 0.337, Q3 = 0.424, max = 0.517).}
\end{figure}

Let us first consider overall performance. Here the most striking result
is that Wikidata configurations outperformed YSO configurations (see
Figure 6, left). However, the explanation of this effect is not evident.
The two main explanatory hypotheses are 1. that the suggestions by the
Wikidata configurations are more salient, or 2. that the native gold
standard is biased towards Wikidata due to its higher coverage of QIDs
as compared to YSO IDs. By contrast, the Wikidata configurations are
more productive than the YSO configurations, and the reason for this
effect is the higher coverage of QIDs as compared to YSO IDs (see Figure
7). By ``productivity'' I mean that the absolute number of assigned
subject terms. So Wikidata configurations are not only qualitatively but
also quantitatively superior to YSO configurations.

@@

Furthermore, the four YSO configurations are more or less on par: the
\texttt{yso-bonsai-en}, \texttt{yso-en}, and \texttt{yso-fasttext-en}
configurations perform very similarly and are slightly outperformed by
\texttt{yso-maui-en} (see Figure 6, left). This result is consistent
with other assessments reporting that ``{[}o{]}f the individual
algorithms, Maui performed best on all corpora'' (Suominen 2019, 13).

A surprising result is that there was no performance gain achieved by
increasing the text basis from titles to titles and abstracts (see
\protect\hyperlink{creating-the-data-foundation}{section ``Creating the
data foundation''}: if anything, the performance for the extended text
basis was slightly worse (see Figure 6, right). There is again no easy
explanation for this find since there are multiple plausible causes
which cannot be distangled here. For example, the abstract of an item in
the Edoc sample data set might be less representative of the item's
content than its title. In that case, the observed worse perfomance of
the Annif configurations based on title and abstracts would even be
expected. Alternatively, it might be the case that the problem lies with
Annif: a larger text basis introduces more confounding factors and might
make it harder to extract the subject terms required. An experimentum
crucis to decide between the two discussed explanations would be to
further extend the text basis to include full texts in addition to
titles and abstracts.

\begin{figure}
\centering
\includegraphics{images/metrics_all_n.pdf}
\caption{Distribution of weighted F1-scores per maximum number of
suggestions 1 \(\leq\) n \(\leq\) 10. See Table 2 for numeric values. n
= 2 is the best perfomring type of configuration. However, it is
outperformed by n = 4 and n = 5 token configurations. The distribution
of the distributions is skewed left. Types of configurations with a
smaller maximum number of suggestions display a more consistent
performance.}
\end{figure}

\begin{table}[]
\begin{tabular}{lllllllllll}
    & n=1   & n=2   & n=3   & n=4   & n=5   & n=6   & n=7   & n=8   & n=9   & n=10  \\ \hline
min & 0.239 & 0.403 & 0.454 & 0.391 & 0.336 & 0.289 & 0.250 & 0.218 & 0.192 & 0.169 \\
Q1  & 0.407 & 0.491 & 0.454 & 0.391 & 0.336 & 0.289 & 0.250 & 0.218 & 0.192 & 0.169 \\
M   & 0.414 & 0.496 & 0.454 & 0.391 & 0.336 & 0.289 & 0.250 & 0.218 & 0.192 & 0.169 \\
Q3  & 0.414 & 0.496 & 0.456 & 0.400 & 0.351 & 0.309 & 0.277 & 0.251 & 0.231 & 0.215 \\
max & 0.414 & 0.496 & 0.492 & 0.517 & 0.503 & 0.467 & 0.427 & 0.389 & 0.355 & 0.324
\end{tabular}
\caption{Distribution of weighted F1-scores per maximum number of suggestions 1 $\leq$ n $\leq$ 10.}
\label{tab:max-number}
\end{table}

Let us turn to the parameter of the maximum number of suggestions per
item in the Edoc sample data set (see Figure 8). Here the best
performance was achieved by a token n = 4 configuration. However, the
type n = 4 configuration performed on average significantly worse than
the type n = 3 and type n = 2 configurations. So with respect to the
maximum number of suggestions, there is a discrepancy between the best
performing token configuration and the best performing type of
configuration. This distinction is important: when choosing a
configuration for the purpose of a production system, we are usually
interested in the best token configuration. The top 20 token
configurations are summarized in Figure 9. The best performing
configuration with a weighted F1-score of 0.517 is \texttt{wikidata-en}
with a maximum number of 4 suggestions per item; the text basis (title
or abstract and title) does not matter.

@@

\hypertarget{results-by-department}{%
\subsubsection{Results by department}\label{results-by-department}}

In \protect\hyperlink{general-results}{section ``General results''} I
have identified and discussed the overall best Annif configerations.
However, in \protect\hyperlink{analysis}{section ``Analysis''}) I had
also noted the caveat that the performance of Annif might vary according
to department due to systematic biases in constructing the Edoc sample
data set. In this section I will therefore assess the performance of the
various Annif configurations per department.

We begin again by constructing the data foundation. Since the Edoc
\texttt{department} data field is mandatory, the sample data set is
partitioned into blocks of departments by default. We can thus create a
data foundation for each such block by calling
\texttt{\_Analysis.super\_make\_metrics} with a departmental parameter:

@@

This yields 1000 files in
\texttt{/metrics/metrics\_\{department\}\_\{marker\}.json} (100
configurations \(*\) 10 departments) where \texttt{department} specifies
the department according to the Edoc convention. The data foundation is
summarized in \texttt{/analysis/metrics.json}.

Let us now look at the results. First consider the distribution of the
performance scores across departments as shown in Figure 11. We can see
that there are striking differences in variability between departments.
The departments with the most consistent weighted F1-scores are Medicine
and Central Services. However, these two departments have also the
lowest maximum F1-scores. It is noteworthy that all other departments
have maximum F1-scores that are well above the F1-score of 0.517 of the
best performing general Annif configuration.

@@

Let us now consider the best performing Annif configurations per
department as summarized in Figure 12.

For each department, we are interested in the highest weighted F1-score
and the corresponding configuration as summarized in Figure X.

@@

@@

\begin{itemize}
\tightlist
\item
  fulltext support
\end{itemize}

@@

\leavevmode\hypertarget{ref-Suominen.2019}{}%
Suominen, Osma. 2019. {``Annif: DIY Automated Subject Indexing Using
Multiple Algorithms.''} \emph{LIBER Quarterly} 29 (1): 1--25.
\url{https://doi.org/10.18352/lq.10285}.
